---
title: "Clustering Weather data of Texas State 2006-2010"
author: "Crupanshu Ashishbhai Udani, Karan Jeeten Thakkar"
date: 08/10/20
output: html_document
---

```{r setup,include = FALSE}
knitr::opts_chunk$set(echo=TRUE,warning = FALSE,message=FALSE)
```
# Importing libraries
```{r}
library(dplyr)
library(readr)
#library(caret)
library(lubridate) # Used for handling date
library(stringr) # Used for String manipulation for feature extractions
library(amap) # Used for k means clustering
library(ggmap)
```

```{bash, engine.opts="-l"}
#awk { gsub(","," ");print>>"hourly_2010.g"} hourly_2010_mod.g
```

# 1 Importing Data and merging into single data frame
```{r Import Data and merging}
stations_location <- read_csv("../data/stations.csv")

columns_names <- c("STN","WBAN","YEARMODA_HR","TEMP","DEWP","DEWP.Count",  "SLP","SLP.Count"  , "STP","STP.Count", "VISIB","VISIB.Count","WDSP","WDSP.Count", "MXSPD",  "GUST", "PRCP"  ,"SNDP" , "FRSHTT")
fileNames_all <- c("./data/datasets_all/hourly_2006.g","./data/datasets_all/hourly_2007.g","./data/datasets_all/hourly_2008.g","./data/datasets_all/hourly_2009.g","./data/datasets_all/hourly_2010_mod.g")
fileNames_789 <- c("../data/datasets_all/hourly_2007.g","../data/datasets_all/hourly_2008.g","../data/datasets_all/hourly_2009.g")
i<-1
for (file in fileNames_789){
  fileData <- read_delim(file,"  ",skip = 1,col_names = columns_names,na =c("", "NA","999.9","9999.9"),col_types ="cicddidididididdcdc-",trim_ws = TRUE) %>% as.data.frame()
  print(dim(fileData))
  if(i==1)
    stations_all <- fileData
  else
    stations_all <- bind_rows(stations_all,fileData)
  i <- i + 1
}
#stations_all <- bind_cols(fileDataList[1],fileDataList[2],fileDataList[3],fileDataList[4],fileDataList[5])
print(dim(stations_all))
head(stations_all)
```

# 2 Filtering out data based on data month and year
```{r}
stations_all %>% mutate(YEAR = year(ymd_h(YEARMODA_HR)),MONTH= month(ymd_h(YEARMODA_HR)),MONTH_DAY=day(ymd_h(YEARMODA_HR)),HOUR=hour(ymd_h(YEARMODA_HR))) ->stations_all
stations_all %>% filter(MONTH==11 & YEAR>2006 & YEAR<2010 ) -> stations_789_nov
summary(stations_789_nov)
#optional (for mem optimization)
rm(stations_all,fileData)
gc()
```


# 2 Preprocessing
```{r TEMP}
#count(filter("TEMP" > 900))
boxplot(stations_789_nov$TEMP,horizontal = TRUE)
#quantile(stations_789_nov$TEMP,na.rm = TRUE)
dens_TEMP<-density(stations_789_nov$TEMP)
plot(dens_TEMP)
# Plotting TEMP less than equal 150 ℉
stations_789_nov%>%select(TEMP)%>%filter(TEMP<=150)%>%table()%>%plot()
# Replace absurdly high values (>150) of TEMP with NA
stations_789_nov$TEMP[stations_789_nov$TEMP > 150] <- NA
# Replace NA values in TEMP with median, grouping rows by station name
impute.median <- function(x) replace(x, is.na(x), median(x, na.rm = TRUE))
stations_789_nov %>% group_by(STN) %>% mutate(TEMP = impute.median(TEMP))
# Plotting TEMP after replacing TEMP > 150 with median
stations_789_nov$TEMP%>%table()%>%plot()
```

```{r PRCP(Not Needed), include = FALSE}
#stations_789_nov$PRCP<-ifelse(stations_789_nov$PRCP=="99.99",NA,stations_789_nov$PRCP)
#PRCP_split<-str_match(stations_789_nov$PRCP,"([0-9.]*)([A-I]*)")
#stations_789_nov$PRCP.Value<-as.double(PRCP_split[,2])
#stations_789_nov$PRCP.Flag<-PRCP_split[,3]
#density(stations_789_nov$PRCP.Value,na.rm = TRUE) %>% plot()
```

```{r WBAN(NOT NEEDED)}
#Remove WBAN as doesn;t provide any information on the weather data
stations_789_nov$WBAN<- NULL
```

# Taking Daily and Monthly average of TEMP,DEWP,STP,WDSP
```{r}
stations_789_nov %>% group_by(STN,YEAR,MONTH_DAY) %>%
  mutate(
      TEMP=mean(TEMP,na.rm = TRUE),
      DEWP=mean(DEWP,na.rm = TRUE),
      STP=mean(STP,na.rm = TRUE),
      WDSP=mean(WDSP,na.rm = TRUE)) %>%
  select(STN,MONTH_DAY,TEMP,DEWP,STP,WDSP) %>% unique() -> stn_789_nov_avged
stn_789_nov_avged %>% ungroup() %>% group_by(STN,YEAR) %>%
  mutate(
      TEMP=mean(TEMP,na.rm = TRUE),
      DEWP=mean(DEWP,na.rm = TRUE),
      STP=mean(STP,na.rm = TRUE),
      WDSP=mean(WDSP,na.rm = TRUE)) %>%
  select(STN,TEMP,DEWP,STP,WDSP) %>% unique() -> stn_789_nov_avged
replace(stn_789_nov_avged$STP,is.nan(stn_789_nov_avged$STP),0) -> stn_789_nov_avged$STP
replace(stn_789_nov_avged$DEWP,is.nan(stn_789_nov_avged$DEWP),0) -> stn_789_nov_avged$DEWP
replace(stn_789_nov_avged$WDSP,is.nan(stn_789_nov_avged$WDSP),0) -> stn_789_nov_avged$WDSP
```

#Merging Location data of station
```{r}
stations_location %>% select(StationNumber,Lat,Lon) %>% group_by(StationNumber) %>%
     mutate(Lat=mean(Lat,na.rm = TRUE),
            Lon=mean(Lon,na.rm = TRUE)) %>% unique() ->stations_location
#merge df for stn loc and stn processed
merge(stn_789_nov_avged,stations_location,by.x="STN",by.y = "StationNumber",all.x = TRUE) -> stn_789_nov_avged_loc
rm(stn_789_nov_avged,stations_789_nov,stations_location)
gc()
```

#Clustering year 2007 k=2 to 8 using Euclidean metric
```{r}
stn_789_nov_avged_loc %>% filter(YEAR==2007) %>% ungroup() %>% select(TEMP,DEWP,STP,WDSP) -> stn_7_nov_avged_loc
maxk<-12
kmn_model<-stn_7_nov_avged_loc %>% Kmeans(x=.,centers = 2,nstart = 2,iter.max = 100,method = "euclidean")
ggplot(stn_789_nov_avged_loc %>% filter(YEAR==2007) %>% select(Lon,Lat) %>% mutate(clustAllocated=as.factor(kmn_model$cluster)) , aes(Lon, Lat,colour=clustAllocated)) +
  geom_point(size = 1, show.legend = TRUE) +
  coord_quickmap()
k<-2
resultMatrix <- data.frame(
Metric=factor(),
K=numeric(),
SSE=numeric(),
YEAR=numeric()
)
while(k <= maxk){
  set.seed(110)
  kmn_model<- stn_7_nov_avged_loc %>% Kmeans(x=.,centers = k,nstart = k,iter.max = 100,method = "euclidean")
  resultMatrix %>% add_row(Metric="euclidean",K=k,SSE=sum(kmn_model$within),YEAR=2007) -> resultMatrix
  k<-k+1
}
```

#Clustering year 2008 k=2 to 8 using Euclidean metric
```{r}
stn_789_nov_avged_loc %>% filter(YEAR==2008) %>% ungroup() %>% select(TEMP,DEWP,STP,WDSP) -> stn_8_nov_avged_loc
k<-2
while(k <= maxk){
  set.seed(110)
  kmn_model<-stn_8_nov_avged_loc %>% Kmeans(x=.,centers = k,nstart = k,iter.max = 100,method = "euclidean")
  resultMatrix %>% add_row(Metric="euclidean",K=k,SSE=sum(kmn_model$within),YEAR=2008) -> resultMatrix
  k<-k+1
}
```

#Clustering year 2009 k=2 to 8 using Euclidean metric
```{r}
stn_789_nov_avged_loc %>% filter(YEAR==2009) %>% ungroup() %>% select(TEMP,DEWP,STP,WDSP) -> stn_9_nov_avged_loc
k<-2
while(k <= maxk){
  set.seed(110)
  kmn_model<-stn_9_nov_avged_loc %>% Kmeans(x=.,centers = k,nstart = k,iter.max = 100,method = "euclidean")
  resultMatrix %>% add_row(Metric="euclidean",K=k,SSE=sum(kmn_model$within),YEAR=2009) -> resultMatrix
  k<-k+1
}
```

# Ploting graph for SSE and K for each year for Euclidean distance metric
```{r}
resultMatrix %>% filter(YEAR==2007) %>% select(K,SSE) %>% plot(type = "b")
resultMatrix %>% filter(YEAR==2008) %>% select(K,SSE) %>% plot(type = "b")
resultMatrix %>% filter(YEAR==2009) %>% select(K,SSE) %>% plot(type = "b")
```
#Clustering year 2007 k=2 to 8 using Pearson metric
```{r}
k<-2
while(k <= maxk){
  set.seed(110)
  kmn_model<-stn_7_nov_avged_loc %>% Kmeans(x=.,centers = k,nstart = k,iter.max = 100,method = "pearson")
  resultMatrix %>% add_row(Metric="pearson",K=k,SSE=sum(kmn_model$within),YEAR=2007) -> resultMatrix
  k<-k+1
}
```

#Clustering year 2008 k=2 to 8 using Manhattan metric
```{r}
k<-2
while(k <= maxk){
  set.seed(110)
  kmn_model<-stn_8_nov_avged_loc %>% Kmeans(x=.,centers = k,nstart = k,iter.max = 100,method = "pearson")
  resultMatrix %>% add_row(Metric="pearson",K=k,SSE=sum(kmn_model$within),YEAR=2008) -> resultMatrix
  k<-k+1
}
```

#Clustering year 2009 k=2 to 8 using Manhattan metric
```{r}
k<-2
while(k <= maxk){
  set.seed(110)
  kmn_model<-stn_9_nov_avged_loc %>% Kmeans(x=.,centers = k,nstart = k,iter.max = 100,method = "pearson")
  resultMatrix %>% add_row(Metric="pearson",K=k,SSE=sum(kmn_model$within),YEAR=2008) -> resultMatrix
  k<-k+1
}
```

# Ploting graph for SSE and K for each year
```{r}
resultMatrix %>% filter(YEAR==2007,Metric=="pearson") %>% select(K,SSE) %>% plot(type = "b")
resultMatrix %>% filter(YEAR==2008,Metric=="pearson") %>% select(K,SSE) %>% plot(type = "b")
resultMatrix %>% filter(YEAR==2009,Metric=="pearson") %>% select(K,SSE) %>% plot(type = "b")
```